Spontaneous Similarity Judgments of Musical Tones played with various Playing Techniques

### 2017-12-05 paper/lagrangeSpontaneousSimilarity2017.tex
xp1

Interestingly, modern physiological models of the primary mamalian auditory system  .  Dau's model acount for frequency and rate of modulation. Shamma's model account for frequency, rate and bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret. Though, most of the numerical experiments done using this model did not fully demonstrated the usefullness of this third dimension.

@@

Thus the second step would be rather opportunistic, in the sense that, from this large set of features, one can pick the ones that are the most relevant for the task at hand.

Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider a notion of dissimilarity and not the

Following .

Spontaneous similarity judgments

@@

\section{Dimensions of Timbre}\label{sec:}

\section{Data}\label{sec:}

@@

@@

Afin de mesurer les performances de l'algorithme, nous avons besoins d'établir une vérité de terrain, en l’occurrence les similarités entre modes de jeu données par des humains.

@@

\section{Modeling}\label{sec:}

@@

>>>>

Interestingly, modern physiological models of the primary mamalian auditory system also considers the evolution of the energy distribution at longer time scales, \textit{i.e.} how the energy modulates through time for given frequencies.  Dau's model acount for frequency and rate of modulation. Shamma's model account for frequency, rate and bandwidth to fully describes the physiological evidence gathered by studying the auditory system of the ferret. Though, most of the numerical experiments done using this model did not fully demonstrated the usefullness of this third dimension.

@@

What the higher level of the auditory cortex does remains largely unkown, but we can assume that the level of plasticity is very high, as any living entity has to adapt their percepts to a wide diversity of tasks.  Considering the lack of information about this stage, we will assume in this paper that this second step is rather opportunistic, in the sense that, from this large set of features, it aims at multiplexing in some way the ones that are the most relevant for the task at hand.

Considering not only the type of musical instrument but also the playing technique is interesting in that matter, as it invite us to consider the relation between the rate of modulation and timbre percepts. A notion that have not been extensively studied in the litterature. It also permits to consider the notion of timbral similarity at finer grain of detail.

@@

@@

The dataset considered in this study has audio recording of 16 different  musical instruments () played with different playing techniques, which leads to 143 different couple of instrument / playing technique. For each couple, the pitch and intonation is varied leading to 25444 samples. The way the dataset is segmented is explained in more details in Appendix \ref{sec:dataset}.

Studies about perceptual similarity are plagued with a dimensional problem. When considering $n$ items, a complete exploration of the similarity space requires the filling of a $n^2$ matrix. Assuming a symmetric similarity (the similarity of A to B is equl to the similarity of B to A) reduces the number but not by much.

In an attempt to reduce the number of items, we decided to perform a selection experiment where the subject is asked to give his opinion on which \ipt is relevant to study. An \ipt is said to be relevant to study if it is likely to be associated to another \ipt. Interest is rated on a 7 ticks scale.

Examples are given :
\begin{itemize}
  \item One star: this \ipt is singular, it is not useful to compare it with another \ipt.
  \item Four stars: there is a proximity on one aspect of the sound between this \ipt and of an \ipt of another instrument, but this one is neither decisive nor obvious.
  \item Seven stars: there is a large similarity between this \ipt and an \ipn of another instrument, it is interesting to keep it.
\end{itemize}

Select which couple (instrument, playing technique) are interesting to consider.
Question: Is it interesting to compare this couple to another couple of a different instrument ?
The question is asked for each couple ()
The subject can listen to all the recordings of each couple

Two musical composition professors of the CNSMDP, a renowned French school of music did the test. All the couples with a grade higher than three are selected, bringing 78 couples to study.

The aim of experiment 1

@@

Afin de mesurer les performances de l'algorithme, nous avons besoin d'établir une vérité de terrain, en l’occurrence les similarités entre modes de jeu données par des humains.

@@

\section{Analysis}\label{sec:}

\section{Model}\label{sec:}

\section{Results}\label{sec:}

\section{Discussion}\label{sec:}

@@

\Appendix

\section{Dataset}
